Grow with the flow: a latitudinal cline in physiology is associated with more variable precipitation in \textit{Erythranthe cardinalis}

### 2017-09-03 ms/ms.Rnw
added effect size figure to ms

<<Effect size confidence intervals, echo=FALSE, results=hide, warning=FALSE, message=FALSE>>=

  # pathFig <- "/Users/cdmuir/Google Drive/CardAdapt/ms/Figures" # for debug

  pdf(paste0(pathFig, "/FigureS_EffectSizes.pdf"), 8, 8, useDingbats = F)
  par(mai = c(1, 1, 0.5, 0.1), mfrow = c(2, 2))

    mod <- mods[[i]]
    main <- c("Leaf expansion", "Stem elongation", "Photosynthesis", "Mortality")[i]
    plot(0, 0, type = "n", axes = FALSE, frame.plot = TRUE,
         ylim = range(mod), xlim = c(1, 7),
         xlab = "", ylab = "Effect size", cex.lab = 1.5,
         main = main)

    abline(h = 0, lty = 2, col = "grey50")
    for (i in 1:7) {
    lines(rep(i, 2), c(mod[i, "2.5 %"], mod[i, "97.5 %"]))
      points(i, mod[i, "estimate"], pch = 19)
    }

    axis(2, las = 1, lwd = 0, lwd.ticks = 1)
    clip(-100, 100, -100, 100)
    text(x = 1:7, y = grconvertY(-0.025, "npc", "user"), srt = 45, pos = 2, 
         labels = c("Pop", "Pop x Temp", "Pop x Water", "Pop x Temp x Water", 
                    "Pop - Pop x Temp", "Pop - Pop x Water", "Pop - Pop x Temp x Water"))

@@

Failure to detect a significant effect could be the result of Type-2 error, so we complemented the above analysis by comparing effect sizes in the full model. The full model contains all main effects, two-way interactions (Population $\times$ Temperature, Population $\times$ Water), a three-way interaction (Population $\times$ Temperature $\times$ Water), and random effects. We compared the effect sizes (mean squared error for LMM, ? for GLMM) of Population to Population $\times$ Treatment in 1000 parametric bootstrap samples simulated from fitted models. MARKER

@@

In contrast to the genetic variation in trait means described above, we found little evidence of G$\times$E in \textit{E. cardinalis}. There were only two statistically significant Population $\times$ Treatment interactions (Table~\ref{table:Table_AnovaSummary}, Fig.~\ref{fig:FigS_RxnNorms}), but these were not strong compared to Population and Temperature effects. Otherwise, populations responded similarly to treatments: faster growth in the hot treatment, slower growth in the dry treatment, and high mortality in the hot, dry treatment (Table~\ref{table:Table_AnovaSummary}). Complete ANOVA tables are available in the Supporting Information (\Cref{table:TableS_fitLER,table:TableS_fitSER,table:TableS_fitPhoto,table:TableS_fitMort})

@@

>>>>

@@

Failure to detect a significant effect could be the result of Type-2 error, so we complemented step-wise ANOVA (see above) by comparing effect sizes calculated in the full model. The full model contains all main effects, two-way interactions (Population $\times$ Temperature, Population $\times$ Water), a three-way interaction (Population $\times$ Temperature $\times$ Water), and random effects. For linear mixed-effects models (leaf expansion, stem elongation, and photosynthesis) we used mean-squared error as a measure of effect size; for GLMM (mortality) we used $\chi^2$ as a measure of effect size. We did not include germination rate because no Population $\times$ Treatment effects were estimated. The difference in effect size of Population versus Population $\times$ Treatment is:

$$ \Delta \textrm{Effect Size}_{\textrm{Pop} - (\textrm{Pop} \times \textrm{Trt})} = \textrm{Effect Size}_\textrm{Pop} - \textrm{Effect Size}_{\textrm{Pop} \times \textrm{Trt}} $$

We calculated $\Delta \textrm{Effect Size}_{\textrm{Pop} - (\textrm{Pop} \times \textrm{Trt})}$ for all two- and three-way Population $\times$ Treatment interactions for each trait. To determine whether $\Delta \textrm{Effect Size}_{\textrm{Pop} - (\textrm{Pop} \times \textrm{Trt})}$ was significantly different than 0, we calculated 95\% confidence intervals using 1000 parametric bootstrap samples simulated from fitted models. If the 95\% confidence interval for a given $\Delta \textrm{Effect Size}_{\textrm{Pop} - (\textrm{Pop} \times \textrm{Trt})}$, was greater than zero then we concluded that the Population effect size was significantly larger than the Population $\times$ Treatment effect size, and vice versa if the confidence internval was less than zero. If the confidence spanned zero, then the effect sizes are not significantly different.

<<Effect sizes, echo=FALSE, results=hide, warning=FALSE, message=FALSE>>=

  pdf(paste0(pathFig, "/FigureS_EffectSizes.pdf"), 8, 8, useDingbats = F)
  par(mfrow = c(2, 2), oma = c(0, 2, 0, 0))

    if (i %in% c(1, 3)) {
      par(mar = c(5, 10, 5, 0))
    } else {
      par(mar = c(5, 0, 5, 10))
    }

    mod <- mods[[i]]
    main <- paste0(LETTERS[i], ". ", 
                   c("Leaf expansion", "Stem elongation", "Photosynthesis", "Mortality")[i])
    plot(0, 0, type = "n", axes = FALSE, frame.plot = TRUE,
         xlim = range(mod), ylim = c(7.5, 0.5),
         ylab = "", xlab = "Effect size", cex.lab = 1.5, main = main)

    abline(v = 0, lty = 2, col = "grey50")
    for (j in 1:7) {
      lines(y = rep(j, 2), c(mod[j, "2.5 %"], mod[j, "97.5 %"]))
      x = points(mod[j, "estimate"], j, pch = 19)
    }

    axis(1, las = 1, lwd = 0, lwd.ticks = 1)
    clip(-100, 100, -100, 100)
   if (i %in% c(1, 3)) {
      text(y = 1:7, x = grconvertX(-0.025, "npc", "user"), srt = 0, pos = 2, 
           labels = c("Pop", "Pop x Temp", "Pop x Water", "Pop x Temp x Water", 
                      "Pop - (Pop x Temp)", "Pop - (Pop x Water)", 
                      "Pop - (Pop x Temp x Water)"))
    }
  }

  # Numbers for Results
  comparisons <- c(
    "PopID - PopID:TempTrt",
    "PopID - PopID:WaterTrt",
    "PopID - PopID:TempTrt:WaterTrt"
  )

  fn <- function(ciMerMod, comparisons) {
    ret <- numeric(3)

    # Pop has larger effect than Pop x Trt
    ret[1] <- length(which(ciMerMod[comparisons, "2.5 %"] > 0))

    # Pop has smaller effect than Pop x Trt
    ret[2] <- length(which(ciMerMod[comparisons, "97.5 %"] < 0))

    # Pop and Pop x Trt have similar effect
    ret[3] <- length(which(ciMerMod[comparisons, "2.5 %"] < 0 & 
                             ciMerMod[comparisons, "97.5 %"] > 0))

    stopifnot(sum(ret) == length(comparisons))

    ret

  effSize <- sapply(list(ciMerModLER, ciMerModSER, ciMerModPhoto, ciMerModMort),
                    fn, comparisons = comparisons)
  colnames(effSize) <- c("LER", "SER", "Photo", "Mort")
  rownames(effSize) <- c("larger", "smaller", "same")

@@

In contrast to the genetic variation in trait means described above, we found little evidence of G$\times$E in \textit{E. cardinalis}. There were only two statistically significant Population $\times$ Treatment interactions (Table~\ref{table:Table_AnovaSummary}, Fig.~\ref{fig:FigS_RxnNorms}), but these were not strong compared to Population and Temperature effects. Otherwise, populations responded similarly to treatments: faster growth in the hot treatment, slower growth in the dry treatment, and high mortality in the hot, dry treatment (Table~\ref{table:Table_AnovaSummary}). Complete ANOVA tables are available in the Supporting Information (\Cref{table:TableS_fitLER,table:TableS_fitSER,table:TableS_fitPhoto,table:TableS_fitMort}).

The effect size of Population was significantly larger than that for Population $\times$ Treatment interactions (Fig.~\ref{fig:FigS_EffectSizes}) in most cases. For leaf expansion, Population had a signicantly larger effect size than Population $\times$ Treatment interactions in \Sexpr{effSize["larger", "LER"]} of \Sexpr{sum(effSize[, "LER"])} comparisons (Fig.~\ref{fig:FigS_EffectSizes}A). For stem elongation (Fig.~\ref{fig:FigS_EffectSizes}B) and mortality (Fig.~\ref{fig:FigS_EffectSizes}D), Population effect sizes were significantly larger than all Population $\times$ Treatment interactions. For Photosynthesis, Population and Population $\times$ Treatment effect sizes were not significantly different (Fig.~\ref{fig:FigS_EffectSizes}C), presumably because we had a smaller sample size.

@@

% Effect sizes of LER, SER, Photo, and Mort

\begin{figure}
	\centerline{\includegraphics{Figures/FigureS_EffectSizes.pdf}}
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[Effect sizes]{Population effect sizes are usually larger than Population $\times$ Treatment effect sizes. In each panel, we plot estimated effect size (points) and 95\% confidence intervals (lines) inferred using parametric bootstrap (see Materials and Methods). At the top, we plot the effect sizes of Population (`Pop'), two-way interactions between Population and Temperature (`Pop $\times$ Temp') or Water (`Pop $\times$ Water'), and the three-way interaction between Population, Temperature, and Water (`Pop $\times$ Temp $\times$ Water'). Below that we plot the difference in the Population minus the Population $\times$ Treatment effect size (e.g. `Pop - (Pop $\times$ Temp)'). When confidence intervals do no overlap zero (dashed line), this means that Population has a significantly greater effect size than the interaction. Effect sizes were measured using unstandarized mean square error for linear mixed-effects models (leaf expansion, stem elongation, and photosynthesis) and $\chi^2$ for GLMM (mortality). Hence, the effect size values are not comparable between different traits.}
	\label{fig:FigS_EffectSizes}
\end{figure}
